`r glue::glue('Compare aggregated and non-aggregated reference: {params$library_id}')`

### 2025-07-07 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/reference-aggregation.Rmd
Apply suggestions from code review
Co-authored-by: Ally Hawkins <54039191+allyhawkins@users.noreply.github.com>
define paths at the top

module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
results_dir <- file.path(module_dir, "results", "singler")
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")

# broad consensus cell type groups
validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"

@@

@@

```{r}
sce_file <- file.path(
  data_dir,
  params$sample_id,
  glue::glue("{params$library_id}_processed.rds")
)

sce_df <- readRDS(sce_file) |>
  scuttle::makePerCellDF() |>
  dplyr::select(
    barcodes,
    consensus_annotation = consensus_celltype_annotation,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  )
```

@@

aggr <- readr::read_rds(
  file.path(
    results_dir,
    params$sample_id,
    "aggregated",
    glue::glue("{params$library_id}_singler-results.rds")
  )
) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes") |>
  dplyr::mutate(ref_type = "aggregated")

non_aggr <- readr::read_rds(
  file.path(
    results_dir,
    params$sample_id,
    "not-aggregated",
    glue::glue("{params$library_id}_singler-results.rds")
  )
) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes") |>
  dplyr::mutate(ref_type = "non-aggregated")
```

@@

## Heatmap: Aggregated vs. consensus annotations

@@

## Heatmap: Non-aggregated vs. consensus annotations

@@

Before plotting cell type labels were harmonized to the extent possible to that `NBAtlas` labels matched the corresponding consensus cell type labels (e.g., `Fibroblast` -> `fibroblast` or `Endothelial` -> `endothelial cell`).
Cell types which not literally just spelling variants were left alone.

>>>>

module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
results_dir <- file.path(module_dir, "results", "singler")
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")
```

```{r url paths}
# broad consensus cell type groups
validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"

@@

```{r file paths}
# SingleR results
aggr_singler_file <- file.path(
  results_dir,
  params$sample_id,
  "aggregated",
  glue::glue("{params$library_id}_singler-results.rds")
)

non_aggr_singler_file <- file.path(
  results_dir,
  params$sample_id,
  "not-aggregated",
  glue::glue("{params$library_id}_singler-results.rds")
)

# SCE file
sce_file <- file.path(
  data_dir,
  params$sample_id,
  glue::glue("{params$library_id}_processed.rds")
)
```

@@

```{r}
sce_df <- readRDS(sce_file) |>
  scuttle::makePerCellDF() |>
  dplyr::select(
    barcodes,
    consensus_annotation = consensus_celltype_annotation,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  )
```

@@

aggr <- readr::read_rds(aggr_singler_file) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

non_aggr <- readr::read_rds(non_aggr_singler_file) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")
```

@@

## Comparison with consensus cell type annotations
### Heatmap: Aggregated vs. consensus annotations

@@

### Heatmap: Non-aggregated vs. consensus annotations

@@

Before plotting cell type labels were harmonized to the extent possible to that `NBAtlas` labels matched the corresponding consensus cell type labels (e.g., `Fibroblast` -> `fibroblast` or `Endothelial` -> `endothelial cell`).
Cell types that were not literally just spelling variants were left alone.
